### 2012-04-25 docs/processing.tex
reshaped directory, clean up, add presentation

\documentclass[a4paper,12pt]{article}
\begin{document}

\section{Iterative Solvers}

To minimize the error produced in the grid search for improperly chosen
$tps$ we find optimal $\alpha$, $\beta$ and crustal thickness $h$ which
define the minimal residual curve. 
\begin{equation}
Tps = h \left( \sqrt{ \frac{1}{\beta^2} - p^2} - \sqrt{ \frac{1}{\alpha^2} - p^2} \right)
\end{equation}

\subsection{Newton's Method}
To find the minimum we seek to minize the sum of the squares of the residual $r=(t_i - f(\mathbf{x}))$ 
where $f(\mathbf{x}) = Tps$ as in \emph{1)} and $\mathbf{x} = \lbrack \alpha,\beta,h \rbrack^T$.
Thus we seek 
\begin{equation}
\mbox{min}\; \vert r^Tr \vert = \mbox{min}\; \vert \phi \vert
\end{equation}
which is true when $\frac{d\phi}{d\mathbf{x}} = g(\mathbf{x}^*) = 0 $. The solution 
$\mathbf{x}^*$ is the initial guess $\mathbf{x}_0 +$ the vector $\mathbf{m} $ which 
finds the minimum $g(\mathbf{x}^*) = 0$. Taking the Taylor approximation gives,
\begin{equation}
g(\mathbf{x}+\mathbf{m}) = g(\mathbf{x}) + \mathbf{Hm} + O(\|mathbf{m}\|)
\end{equation}
The hessian $\mathbf{H}$ is equal to $\nabla^2\phi$ which is given by
$\sum_{i,j=1}^{n} \frac{\partial^2 \phi}{\partial x_i \partial x_j}$. As 
$g(\mathbf{x}+\mathbf{m})=0$ we can rewrite \emph{3)} as
\begin{equation}
\mathbf{m}=-\mathbf{H}^-1g(mathbf{x})
\end{equation}
and iterate until convergence with
\begin{equation}
\mathbf{x}_{k+1} = \mathbf{x}_k + \mathbf{m}
\end{equation}

\end{document}